---
title: "Reproducible Research: Peer Assessment 1"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='figure/',
                      echo=FALSE)
```


## Loading and preprocessing the data
##### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
activity <- read.csv2('activity.csv', sep = ',', stringsAsFactors = FALSE)
```
##### 2. Process/transform the data (if necessary) into a format suitable for your analysis

-----
  
## What is mean total number of steps taken per day?

##### 1. Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
steps <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
qplot(steps, xlab='Total number of steps per day', ylab='Frequency', binwidth=500)
```

##### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
stepsmean <- mean(steps)
stepsmedian <- median(steps)
```
We get:

Mean: `r stepsmean`

Median:  `r stepsmedian`

-----
  
## What is the average daily activity pattern?
```{r}
average <- aggregate(x=list(mean=activity$steps), by=list(interval=activity$interval), FUN=mean, na.rm=TRUE)
```

##### 1. Make a time series plot
```{r}
ggplot(data=average, aes(x=interval, y=mean)) +
  geom_line() +
  xlab("5-min interval (starting time)") +
  ylab("average number of steps taken") +
  theme_bw()
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
max <- which.max(average$mean)
time <- average$interval[max]
```

Most steps taken during the 5 minute interval starting at time: `r time`

----
  
## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 
```{r}
missing <- length(which(is.na(activity$steps)))
```

Number of missing values: `r missing`

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
We shall replace with missing values with the mean of the steps taken over all days.

##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activityimputed <- activity
activityimputed$steps[is.na(activityimputed$steps)] <- mean(activity$steps, na.rm=TRUE)
```


##### 4. Make a histogram of the total number of steps taken each day 
```{r}
stepsimputed <- tapply(activityimputed$steps, activityimputed$date, sum)
qplot(stepsimputed, xlab='Total number of steps per day', ylab='Frequency', binwidth=500)
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
stepsmean <- as.integer(mean(stepsimputed))
stepsmedian <- as.integer(median(stepsimputed))
```
Mean: `r stepsmean`

Median:  `r stepsmedian`


----
  
## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
  
```{r}
activityimputed$daytype <-  ifelse(as.POSIXlt(activityimputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot

```{r}
averageactivityimputed <- aggregate(steps ~ interval + daytype, data=activityimputed, mean)
ggplot(averageactivityimputed, aes(interval, steps)) + 
  geom_line() + 
  facet_grid(daytype ~ .) +
  xlab("5 min - Interval (starting time)") + 
  ylab("avarage number of steps")+
  theme_bw()
```

Yes there are differences, people get out of bed later during weekends, and move more during the day(presumably they sit still more during working hours).